NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE, PG and PS lipids

### 2017-08-23 Manuscript/manuscript.tex
Add experimental results for the signs of PS headgroup order parameters

\subsection{PS headgroup}
Experimental order parameters for bilayers with PS headgroups are 
available in literature only for DOPS  \cite{browning80}. These are shown together
with the results from different simulation models in Fig. \ref{HGorderParametersDOPS}.
The signs are not yet know experimentally.
Before the experimental data becomes available we have denoted 
negative signs for $\beta$ and $\alpha$ segment order parameters,
because this selection of signs is in best agreement with the most simulation
results. With this interpretation the CHARMM36, CHARMM36ua and Slipids
models would be in relatively good agreement with experimental data in
the headgroup region. However, if the experimental signs turn out to
be different than expected, this conclusion will change.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersDOPS.eps}
  \caption{\label{HGorderParametersDOPS}
    Order parameters for DOPS headgroup and glycerol
    backbone from simulations with different models and experiments without CaCl$_2$ 
    Experimental data from \cite{browning80} contains 0.1M of NaCl.
    Signs are not known experimentally. They chosen to give best agreement simulations.
  }
  \todo{Experimental signs of the order parameters would highly useful.} \\
\end{figure}

The glycerol backbone order parameters for Slipids model are, however,
far from experimental values, as previously observed also for PC 
lipids \cite{botan15}. The glycerol backbone structures between PC and
PS lipids simulated with CHARMM36 are compared with the structures
simulated with CHARMM36 in Fig. \ref{glycerol_buslaev}.
The differences in sampled conformation leading to the order parameter
differences are clearly visible in the figure.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/glycerol_buslaev.png}
  \caption{\label{glycerol_buslaev}
    Snapshots overlayed from different simulations for glycerol backbone region
    by Pavel Buslaev.
  }
\end{figure}

%Comparison between experiments and simulations with CHARMM GUI
%parameters for PS lipids in Fig. \ref{HGorderParametersPOPS} suggest that
%the model overestimates magnitudes in $\beta$ and g$_3$ order parameters
%and forking in $\alpha$.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPOPS.eps}
  \caption{\label{HGorderParametersPOPS}
    Order parameters for POPS (top) and DOPS (bottom) headgroup and glycerol
    backbone from simulations with different models and experiments without CaCl$_2$ 
    (DOPS from \cite{browning80} contains 0.1M of NaCl, POPC:POPS mixture from \cite{roux90}).
    Absolute values are shown for experimental data, because signs are not known.
    Simulations values are -S$_{CH}$
  }
  \todo{More simulation data for lipids with different headgroups to be collected} \\
  \todo{Some simulations contain potassium as counterions, while some sodium.
    All experiments here contain some amount of sodium salt. The best
    ion concentrations for comparison should be figured out.} \\
  \todo{Experimental signs of the order parameters would highly useful.} \\
  \todo{Why there is difference between CHARMM36 simulation results from POPS:POPC mixture and
  pure POPS? Discussion in https://github.com/NMRLipids/NMRlipidsIVotherHGs/issues/1}
\end{figure}

@@

>>>>

\subsection{PS headgroup}
Order parameters for bilayers with PS headgroups
from different experiments and simulations are shown in
Figs. \ref{HGorderParametersDOPS} and  \ref{HGorderParametersPOPS}.
None of the tested models gives satisfactory agreement with
experiments for order parameters in headgroup $\alpha$ and $\beta$
carbons.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersDOPS.eps}
  \caption{\label{HGorderParametersDOPS}
    Order parameters for DOPS headgroup and glycerol
    backbone from simulations with different models and experiments without CaCl$_2$ 
    Experimental data from \cite{browning80} contains 0.1M of NaCl.
    Signs are taken from experiments for POPS described in Supplementary Information.
  }
\end{figure}
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPOPS.eps}
  \caption{\label{HGorderParametersPOPS}
    Headgroup and glycerol backbone order parameters for systems with POPS 
    from simulations with different models and experiments \cite{roux90}.
    Signs are taken from experiments described in Supplementary Information.
  }
  \todo{More simulation data for lipids with different headgroups to be collected} \\
  \todo{Some simulations contain potassium as counterions, while some sodium.
    All experiments here contain some amount of sodium salt. The best
    ion concentrations for comparison should be figured out.} \\
  \todo{Why there is difference between CHARMM36 simulation results from POPS:POPC mixture and
  pure POPS? Discussion in https://github.com/NMRLipids/NMRlipidsIVotherHGs/issues/1}
\end{figure}

Glycerol backbone order parameters seems similar in all models,
except in Slipids. Even thought glycerol backbone order parameter
values are not yet experimentally available for PS lipids,
the comparison with the results for PC lipids suggest that
Slipid model do not correctly capture the glycerol backbone
structure \cite{botan15}. The glycerol backbone structures between PC and
PS lipids simulated with CHARMM36 are compared with the structures
simulated with CHARMM36 in Fig. \ref{glycerol_buslaev}.
The differences in sampled conformation leading to the order parameter
differences are clearly visible in the figure.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/glycerol_buslaev.png}
  \caption{\label{glycerol_buslaev}
    Snapshots overlayed from different simulations for glycerol backbone region
    by Pavel Buslaev.
  }
\end{figure}

%Comparison between experiments and simulations with CHARMM GUI
%parameters for PS lipids in Fig. \ref{HGorderParametersPOPS} suggest that
%the model overestimates magnitudes in $\beta$ and g$_3$ order parameters
%and forking in $\alpha$.

@@

\section{Measurements of order parameter sign}

Fig. \ref{PShgSIGNS} summarizes the experimental results on the order parameter sign
measurement for POPS sample. The experimental protocol is the same used in Ref. \citenum{ferreira16}.
In (a) you see the headgroup region of the INEPT spectrum where alpha and beta are
identified. In (b) you have the R-PDLF slices for alpha and beta where you see one single
splitting for beta (which gives an order parameter equal to 0.12), and for alpha a superposition
of a large splitting (order parameter equal to 0.09) and a very small splitting which cannot be
calculated. On the bottom you have the S-DROSS slices of these two carbons. The grey lines show a
random collection of slices from noise such that it gets clear what is significant. The S-DROSS
slice for beta clearly shows that the order parameter is negative. The slice for alpha shows that
the higher order parameter is positive and suggests that the smaller order parameter is negative
(from the deviation towards negative values in the longer t1 times).
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/PShgSIGNS.pdf}
  \caption{\label{PShgSIGNS}
    Experimental results for sign measurement for POPS sample
  }
\end{figure}

The results updated with SIMPSON simulations for the SDROSS profiles
are shown in Fig. \ref{PShgSIGNSsimpson}. The value for the smaller
alpha order parameter is taken from Fig 3 in Ref. \citenum{roux91},
because resolution in 13C NMR experiments was nor high enough to determine
numerical value for this. The plots in Fig. \ref{PShgSIGNSsimpson} (c) show
the following. The error bars and points are the experimental SDROSS data.
The thick lines are SIMPSON simulations. The simulations were done by using
the order parameter for beta equal to -0.12 and for alpha one order parameter
equal to 0.09 and the other equal to -0.02 (black) or 0.02 (grey).
Since the black lines agree with experimental data, we conlude that
the order parameters for $\beta$ carbon are -0.12 and for $\alpha$
order parameters are 0.09 and -0.02.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/PShgSIGNSsimpson.pdf}
  \caption{\label{PShgSIGNSsimpson}
    Experimental results for sign measurement for POPS sample
  }
\end{figure}
